Behavior Driven Development in Haskell

### 2012-05-25 paper.tex
brushing up.

\keywords Haddock, QuickCheck, doctest, hspec

@@

In this paper, we consider how to put BDD into practice in Haskell.
We brought BDD methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}.
We also implemented {\tt hspec} inspired by rspec~\cite{rspec}
and integrated with QuickCheck.

@@

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
\item Introduced {\tt hspec} from Ruby community to Haskell community and
integrated it with QuickCheck
(Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
\item Show best current practice of behavior driven development in Haskell (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use {\tt hspec} aslo integrated with QuickCheck.

@@

\noindent Since the right hand side of {\tt ->} is verbose as documentation,
open lambda terms are also allowed:

@@

@@

@@

We implemented the {\tt >>>} syntax and it
is available in Haddock version 2.8 or later.
We also extended Haddock to support the {\tt prop>} syntax
which will be available in the future release.

@@

Since no-parameter properties have no parameters,
it is enough to test them just one.
However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
More precisely, the evaluation results are memorized and they are
compared with {\tt True} 100 times.
The current maintainer of QuickCheck modified that
no-parameter properties are tested just once in its source repository.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{hspec}
\label{sec:hspec}

@@

@@

\noindent We should execute {\tt doctest} to ensure no syntax errors.
Also, we specifies internal behaviors with {\tt hspec}:

@@

\noindent Also, we should execute this hspec module
to ensure no syntax errors.
Then we implement the function.

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as ghc-mod) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock annotations.

@@

@@

\item @kazu: Can you add a reference to:
\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
\item @kazu: should this subsection go into "best current practice"?

>>>>

\keywords Haddock, QuickCheck, doctest, Hspec

@@

In this paper, we consider how to put BDD into practice in Haskell.
We brought BDD methods from other programming language communities and
integrated them with a Haskell-original test method.
In particular, we introduced {\tt doctest}~\cite{doctest} and
integrated with QuickCheck~\cite{quickcheck}.
We also implemented {\tt Hspec} inspired by rspec~\cite{rspec}
and integrated with QuickCheck.

@@

\begin{itemize}
\item Introduced {\tt doctest} from Python community to Haskell community
and integrated it with QuickCheck
(Section \ref{sec:Visible-behavior} and Section \ref{sec:doctest})
\item Introduced {\tt Hspec} from Ruby community to Haskell community and
integrated it with QuickCheck
(Section \ref{sec:Internal-behavior} and Section \ref{sec:hspec})
\item Show best current practice of behavior driven development in Haskell (Section \ref{ref:Best-current-practice})
\end{itemize}

@@

To specify visible behavior, we propose to add two new annotation syntax to Haddock and to use {\tt doctest} integrated with QuickCheck. To describe internal behavior, we propose to use {\tt Hspec} aslo integrated with QuickCheck.

@@

\noindent Since the right hand side of {\tt ->} is verbose as documentation,
lambda terms without the lambda prefix are also allowed:

@@

\begin{quote}
    \emph{TBD: provide an example?}
\end{quote}

@@

\begin{quote}
    \emph{TBD: should user be replaced with developer?}
\end{quote}

@@

We implemented the {\tt >>>} syntax and it
is available in Haddock version 2.8 or later.
We also extended Haddock to support the {\tt prop>} syntax
which will be available in the future release.
Both examples and properties in Haddock annotation are pretty printed when producing document
such as in HTML.

@@

Since no-parameter properties have no parameters,
it is enough to test them just one.
However, QuickCheck 2.3 or earlier tests no-parameter properties 100 times.
More precisely, the evaluation results are memorized and they are
compared with {\tt True} 100 times.
The current maintainer of QuickCheck modified so that
no-parameter properties are tested just once in its source repository.
This feature will be available in the future release.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Hspec}
\label{sec:hspec}

@@

\begin{quote}
    \emph{TBD: one more example relating to IO is desired.}
\end{quote}

@@

\noindent We should execute {\tt doctest} to ensure no syntax errors.
Also, we specifies internal behaviors with {\tt Hspec}:

@@

\noindent Also, we should execute this Hspec module
to ensure no syntax errors.
Then we implement the function.

@@

Note that the authors are not fundamentalist.
If functions to be defined are simple enough,
we might implement them first withtout signatures.
And their signatures can be automatically
inserted by IDE (such as Emacs/vim with {\tt ghc-mod}) later
if Haskell compilers can infer correctly.
Also, we can cut and paste examples and their results
from Haskell interpreters to their Haddock annotations.

@@

\begin{quote}
\emph{How about moving this to appendix?}
\end{quote}

@@

\item @kazu: Can you add a reference to:
\url{http://www.haskell.org/cabal/users-guide/developing-packages.html#test-suites}
